The computational capacity of graphics cards for general-purpose computing have been progressing fast over the last decade. A major reason is demanding computer games, where standard of performance and high quality graphics are constantly raised. Another reason are better suitable technologies for programming the graphics cards. Combined, the product are high raw performance devices and means to access that performance. This thesis investigates some of the current technologies for general-purpose computing on graphics processing units. Technologies are primarily compared by means of benchmarking performance and secondarily by factors concerning programming and implementation. The selection of technology can have a large impact on performance. The benchmark application found the difference of the fastest technology, CUDA, compared to the slowest, OpenCL, to be twice as long execution time. The application also found out that the older technologies, OpenGL and DirectX, are competitive with CUDA and OpenCL in terms of the resulting performance speed.%151
%
%En sammanfattning ska kort och koncist beskriva och motivera det studerade problemet, metoden samt resultat och slutsatser. Arbetets bidrag till huvudområdet ska tydligt framgå. Vad är det rapporten säger om huvudområdet som vi inte visste tidigare? Exempel på bidrag kan vara vilken effekt en specifik algoritm eller programutvecklingsmetod får i en specifik tillämpning.
%
%Normalt ska en sammanfattning vara högst 150 ord, och inte innehålla några referenser eller radbrytningar.
%
%Sammanfattning på svenska såväl som engelska (abstract) måste finnas med. Om rapporten är skriven på engelska räcker det med engelsk sammanfattning 